\begin{abstract}
    This thesis addresses the fundamental incompatibility of applying gradient
    projection-based continual learning (CL) methods to hypernetwork architectures,
    establishing a novel framework for stable subspace optimization.  Traditional
    projection methods collapse due to a structural–numerical pathology where the
    architectural necessity of weight-chunking, required to avoid an output parameter
    explosion, forces the model to be produced in hundreds or thousands of sequential
    chunks before a single projection step can execute.  This repetitive chunk-generation
    loop accumulates an exceptionally large gradient volume within the memory matrix,
    blowing up the condition number of the projection correlation matrix to infinity
    and causing catastrophic inversion failures that the standard regularization
    parameter $\lambda$ cannot rescue.
    
    The primary research objective is to eliminate these numerical singularities and
    systematically evaluate whether combining hard geometric projections on top of
    task-conditioned hypernetworks and their output-scoped functional regularizers
    yields a constructive synergy.  To achieve this, we introduce
    \emph{projection-scoped normalization}: gradient columns stored in the memory
    matrix $G$ are rescaled to unit $\ell_2$ norm, and the diagonal empirical Fisher
    vector is rescaled by its absolute maximum, together stabilizing the metric tensor
    within the local projection algebra.  We also introduce \texttt{iFOPNG}, an inertial
    variant of Fisher-Orthogonal Projected Natural Gradient Descent \citep{garg_fisher-orthogonal_2026}
    that replaces the current-task Fisher $F_{\text{new}}$ with the combined curvature
    $F_c = F_{\text{new}}+F_{\text{old}}$, inspired by the parameter-inertia concept
    of Elastic Weight Consolidation.  Methods are benchmarked on sequential Split and Permuted MNIST, Split-CIFAR10 and Split-CIFAR100.  The empirical findings reveal that projection normalization
    completely resolves matrix inversion collapse, and that in the hypernetwork setting
    all normalized projection methods outperform plain SGD, with \texttt{iFOPNG} consistently outperforming \texttt{FOPNG} 
    across all benchmarks.  However, a standalone hypernetwork optimized
    with \texttt{Adam} and von Oswald's functional regularizer consistently surpasses
    all projection variants, indicating that momentum, which projection methods forgo by
    design, is the decisive factor in compressed generative weight spaces.
\end{abstract}