\section{Conclusion}
\label{sec:conclusion}

This thesis investigated the extent to which gradient projection methods
can be successfully integrated into chunked hypernetwork architectures for
continual learning, and identified the structural conditions under which
that integration yields benefit or harm.
The primary research question decomposed into four sub-questions addressing
constructive synergy, gradient normalization, the elastic Fisher metric, and
the Fisher accumulation operator.

\paragraph{Sub-RQ1 — Partial synergy bounded by the momentum problem.}
The integration is constructive but incomplete.
No projection variant matches a standalone regularized hypernetwork
optimized via \texttt{Adam}, which leads by approximately $4.5$~pp on
Split-CIFAR10 and by $20$--$40$~pp in the extreme compression regime
($d_h = 8$).
The primary obstacle is structural: projection methods modify the raw
gradient before the optimizer state update, effectively discarding the
first-order momentum that stabilizes generative manifold navigation in
compressed weight spaces.
This interaction has been independently identified as problematic under
certain curvature regimes \citep{hu_hidden_2026}, and resolving it will
require projection-compatible adaptive optimizers rather than post-hoc
gradient modification.

\paragraph{Sub-RQ2 — Normalization is a necessary condition.}
Without explicit local rescaling, projection methods fail catastrophically
on chunked hypernetworks: the condition number of the projection kernel
reached $10^{10.8}$, rendering matrix inversion numerically unreliable.
QR orthonormalization of the gradient basis combined with absolute-maximum
scaling of the metric tensor $F_c$ reduces the condition number by more
than seven orders of magnitude — bounding it within
$[10^{1.3},\,10^{3.3}]$ — and fully recovers projection performance
across all three seeds on the Split-CIFAR10 hypernetwork.
This normalization scheme is a necessary engineering precondition for any
future combination of gradient projection with chunked meta-learners,
regardless of which projection algebra is employed.

\paragraph{Sub-RQ3 — Parameter inertia improves retention on high-overlap benchmarks.}
Replacing \texttt{FOPNG}'s task-specific Fisher metric with the combined
metric $F_c = \hat{F}_\mathrm{new} + \hat{F}_\mathrm{old}$ consistently
improves task retention when inter-task Fisher overlap is substantial.
The advantage of \texttt{iFOPNG} over \texttt{FOPNG} grows monotonically
with benchmark difficulty: negligible on Permuted-MNIST
($+0.1$~pp), substantial on Split-CIFAR10 multi-head ($+14.1$~pp),
and meaningful in the hypernetwork settings ($+9.8$~pp and $+8.6$~pp).
\texttt{iFOPNG} thereby provides a principled and practically effective
improvement over \texttt{FOPNG} at no additional hyperparameter cost.

\paragraph{Sub-RQ4 — MAX accumulation is the clear practical choice.}
On a twenty-task Permuted-MNIST sequence evaluated across five seeds, the
element-wise maximum (MAX) and exponential moving average (EMA)
accumulation strategies produced nearly parallel accuracy and backward
transfer trajectories throughout the full sequence.
A paired $t$-test confirms that MAX is the superior operator:
$t(4) = 14.43$, $p < 0.001$, Cohen's $d = 7.22$, with MAX leading by a
consistent $1.51$--$1.65$~pp across every seed pair.
The effect is systematic and statistically unambiguous, though small in
absolute terms: $1.60$~pp on a benchmark where both strategies decline by
more than $30$~pp over the full sequence --- practically negligible in
deployment terms.
The MAX operator is therefore preferred as the simpler, marginally better,
and hyperparameter-free alternative.

\paragraph{Implications for the field.}
The results establish a clear empirical boundary for the current generation
of gradient projection methods: they can be made numerically stable in
chunked hypernetwork settings via local rescaling, and their Fisher geometry
can be improved via elastic accumulation, but the momentum incompatibility
prevents them from reaching the performance frontier of adaptive
unconstrained optimizers in generative parameter spaces.
The normalization contribution is directly applicable to any method that
maintains a gradient subspace memory alongside a chunked weight generator.
The iFOPNG contribution extends naturally to any continual learning
setting where Fisher overlap accumulates across tasks.
Together, they narrow the gap between projection-based and
regularization-based approaches, and provide a principled foundation for
the next step: a projected adaptive optimizer that respects both the
geometric constraints of the projection and the curvature structure that
adaptive moment estimates capture.

\paragraph{Future directions.}
Three directions are most directly motivated by the findings of this thesis.
First, a projected variant of Adam that maintains momentum across projected
steps — rather than discarding it at each update — would address the
primary mechanism identified in Sub-RQ1 and has been partially explored in
concurrent work \citep{hu_hidden_2026}.
Second, the natural gradient preconditioner for chunk embeddings under the
pullback metric $G_i = J_i^\top J_i$ was implemented and found to interact
destructively with the FOPNG projection when applied to the raw gradient;
applying it to the projected update $v^*$ instead would resolve the metric
mismatch and merits empirical evaluation.
Third, extending the evaluation to longer task sequences and larger
hypernetwork capacities would test whether the partial synergy identified
here scales favourably as the projection subspace budget grows relative to
the task manifold — the regime in which hard constraints are expected, by
theory, to provide the greatest benefit over soft regularization.