\section{Discussion}
\label{sec:discussion}

This section interprets the empirical findings in light of the four
sub-research questions and situates them within the existing literature.
The overarching goal was to determine whether gradient projection methods
can be fruitfully integrated into chunked hypernetworks for continual
learning, and to identify the structural conditions under which this
integration succeeds or fails.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Sub-RQ1: Constructive Synergy or Architectural Conflict?}

Projection methods do not outperform a well-tuned regularized
hypernetwork optimized via \texttt{Adam}.
The gap of $4.5$~pp at task~5 in favor of \texttt{Adam} is consistent
across seeds and widens substantially in the suffocated regime ($d_h = 8$,
Split-MNIST SH), where it reaches $20$--$40$~pp.
We attribute this to a fundamental architectural incompatibility.
Projection methods operate by modifying the raw gradient before the
optimizer state update, which means the first-order momentum estimate
and the adaptive scaling factor maintained by Adam are effectively
discarded at each projected step.
In the highly nonlinear, compressed weight space of a hypernetwork
meta-generator, momentum is not merely a convergence accelerant — it
is a structural regularizer that prevents the optimizer from committing
prematurely to poorly-conditioned gradient directions.
Re-initializing the optimizer's internal state at each step destabilizes
the generative manifold and reduces the quality of the generated weights,
particularly under extreme bottleneck compression.
This finding is consistent with \citet{hu_hidden_2026}, who demonstrate
that gradient modification methods interact adversely with Adam's internal
state under certain curvature regimes, and aligns with the broader
observation that projection-based continual learning methods were
originally designed for SGD-like updates.

The ten-task Split-CIFAR100 result introduces a partial rank reversal
that complicates the Sub-RQ1 conclusion: \texttt{EWC} unexpectedly leads
at $45.5\%$, overtaking both \texttt{Adam} ($42.0\%$) and \texttt{iFOPNG}
($37.8\%$).
This inversion suggests that at higher task counts, soft functional
regularization in output space may dominate hard gradient projection
because the projection memory cannot hold sufficient historical directions
within a fixed column budget.
As the number of tasks grows, the gradient subspace spanned by $G$ becomes
an increasingly coarse approximation to the full historical interference
manifold, while the functional regularizer in output space grows more
informative with each additional stored weight snapshot.
Whether this effect persists at the twenty-task scale, and whether an
expanded memory budget would recover the projection advantage, remains to
be investigated.

\paragraph{First-task initialization and gradient contamination.}
The comparison between \texttt{AdamW} and \texttt{Adam} first-task
initialization confirmed the contamination hypothesis for gradient
projection methods in the hypernetwork setting.
Standard \texttt{Adam} couples weight decay into the gradient estimate as
an additive $\ell_2$ term, so both the gradient memory $G$ and the Fisher
diagonal $\hat{F}$ collected at the end of task~1 inherit a component
aligned with the weight magnitude vector rather than the task loss
curvature.
This contaminates the projection subspace: downstream tasks are projected
away from directions spuriously marked as historical but actually reflecting
optimizer regularization rather than task geometry.
\texttt{AdamW} initialization, which applies weight decay as a decoupled
parameter-shrinkage step, consistently produced lower first-task trajectory
variance and a cleaner initial projection basis.
This finding carries a practically important implication for any system
bootstrapping a projection memory from first-task gradients: the
decoupled optimizer should be preferred at that stage regardless of which
optimizer is used in subsequent tasks.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Sub-RQ2: Projection-Scoped Normalization Resolves the Pathology}

This result contributes a practical engineering solution that is
necessary for any future work combining hard projection constraints with
chunked meta-learners, regardless of which projection algebra is employed.
The normalization is applied locally within the projection computation and
does not alter the underlying optimization landscape: subspace membership
is invariant to positive scalar rescaling of basis vectors, so neither
operation changes the geometry of the projection.
The fundamental problem is therefore purely numerical rather than
conceptual: the intersection of chunked hypernetworks and gradient
projection is geometrically coherent but arithmetically fragile without
explicit local rescaling.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Sub-RQ3: Parameter Inertia via the Elastic Fisher Preconditioner}

The \texttt{iFOPNG} advantage scales monotonically with benchmark difficulty
across all six settings (Figure~\ref{fig:ifopng_comparison}): negligible where
inter-task Fisher overlap is low, substantial on high-overlap benchmarks.
$F_c$ inflates the effective curvature of historically important coordinates,
suppressing over-aggressive natural gradient steps without any penalty term or
stored reference point.

This mechanism stands in direct contrast to EWC, which achieves a similar
dampening effect via an explicit quadratic penalty pulling parameters toward
a reference point.
\texttt{iFOPNG} applies no such restoring force and maintains no reference:
a parameter that mattered before is not pulled back, it is made harder to
displace.
The empirical advantage of \texttt{iFOPNG} over \texttt{FOPNG} on
high-overlap benchmarks confirms that the rigidity of the pure projection
boundary is a genuine limitation of FOPNG, and that lifting it via the
combined metric $F_c$ is a principled and effective remedy.
Nevertheless, \texttt{iFOPNG} does not close the gap to \texttt{Adam},
confirming that the elastic preconditioner addresses rigidity but cannot
recover the momentum mechanism that \texttt{Adam} leverages for efficient
navigation of the generative manifold.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Sub-RQ4: Fisher Accumulation Operator}

This null result is itself informative.
The theoretical argument for EMA was that monotonic MAX growth would
impose a steadily increasing effective mass on parameters that were
important to any prior task, eventually over-constraining the natural
gradient and harming plasticity on new tasks.
The absence of this effect on Permuted-MNIST suggests that, under the
normalization scheme applied here (which rescales both Fisher terms by the
current maximum of $F_c$), the absolute scale of $F_\mathrm{old}$ does not
enter the projection computation — only its relative distribution does.
The max-normalized Fisher is therefore insensitive to whether the
accumulation operator is decaying or non-decaying, as long as the
distribution of importance across parameters is preserved.

Two limitations of this finding should be noted.
First, Permuted-MNIST is a structurally homogeneous benchmark in which
all tasks share the same label space and architecture; inter-task Fisher
overlap is low and approximately uniform across tasks.
On benchmarks with heterogeneous task distributions and non-uniform
overlap, the two operators may diverge as the historical Fisher accumulates
disproportionate mass in task-specific parameter subspaces.
Second, the experiment used a fixed EMA decay of $\alpha = 0.5$;
an adaptive schedule that increases decay as the task sequence
grows longer may produce a more favorable comparison for EMA under
long-horizon retention.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Limitations}

Several limitations should be acknowledged.

First, the hypernetwork results presented here are restricted to
Split-MNIST and Split-CIFAR10 with five tasks.
Longer task sequences may produce different relative orderings:
in particular, the advantage of hard projection constraints over soft
functional regularization may grow with sequence length, as soft
penalties are known to suffer from compounding output drift
\citep{farquhar_towards_2019}, while the projection subspace scales with
the gradient budget rather than the task count.
The preliminary ten-task Split-CIFAR100 result, in which EWC overtakes both
\texttt{Adam} and \texttt{iFOPNG}, is suggestive but insufficient to
establish a trend; a systematic sweep over task count is warranted.

Second, the hypernetwork architecture used in this study is intentionally
shallow, with small hidden dimensions ($d_h \in \{8, 16\}$) and
embedding dimensions of $32$.
Larger hypernetworks, such as those used for language model weight
generation \citep{chauhan_brief_2024}, may exhibit qualitatively different
gradient accumulation dynamics — both in terms of the severity of the
normalization pathology and in the relative contribution of momentum to
meta-generator stability.
The suffocated regime ($d_h = 8$) explored here provides an empirical
lower bound on expressivity, but an upper bound in a high-capacity regime
remains uncharted.

Third, the proposed normalization scheme — QR orthonormalization of the
gradient basis combined with absolute-maximum scaling of the metric tensor
$F_c$ — is heuristic in nature.
It discards the eigenvalue structure of the pullback metric and applies
isotropic rather than curvature-aware rescaling to the chunk embedding
directions.
The principled alternative is the Riemannian gradient under the pullback
metric $G_i = J_i^\top J_i$, where $J_i = \partial W_i / \partial c_i$
is the Jacobian of the chunk-generating map:
\begin{equation}
    \Delta c_i = -\eta\,(J_i^\top J_i + \lambda I)^{-1}
                 \nabla_{c_i}\mathcal{L}.
\end{equation}
This preconditioner was implemented using batched forward-mode automatic
differentiation across all $K$ chunks simultaneously.
When applied to the raw gradient \emph{prior} to the FOPNG projection
step, however, it consistently degraded continual learning performance in
preliminary evaluation: backward transfer worsened by approximately $0.33$
units on Split-MNIST SH relative to the heuristic baseline.
We attribute this to a metric mismatch: the FOPNG projection operator
is calibrated for gradients in Euclidean parameter space, and
preconditioning only the chunk embedding component before projection
produces a geometrically inconsistent input to the kernel
$G^\top \hat{F} G$.
The principled resolution — applying the natural gradient to the
projected update $v^*$ rather than the raw gradient, so that the
projection sees Euclidean inputs and the Riemannian step follows — is left
to future work.