\section{Data Source, Ethics, Code, and Technology Statement}
\label{sec:dsect}

\noindent\textbf{Data Source.} The MNIST (owned by LeCun et al.) and CIFAR (CIFAR-10 and CIFAR-100, owned by Krizhevsky et al.) datasets are public, anonymized datasets licensed for research. These datasets form the basis for the Split-MNIST, Permuted MNIST, Split-CIFAR10, and Split-CIFAR100 benchmarks evaluated in this work. No human or animal data was collected. Data was acquired via the \texttt{torchvision.datasets} API.

\paragraph{Figures.} All figures and plots were created entirely by the author using \texttt{matplotlib} and Weights \& Biases.

\paragraph{Code.} The baseline FOPNG framework was adapted from \citet{garg_fisher-orthogonal_2026}; all other execution infrastructure was implemented by the author (\url{https://github.com/kelares/HyperFisher}). Libraries used: Python (v3.10), PyTorch (v2.x), Torchvision, Weights \& Biases, NumPy, Matplotlib, and Tqdm.

\paragraph{Technology.} The manuscript was typeset in \LaTeX{} using BibTeX. During the preparation of this work, the author(s) used Gemini and Claude in order to optimize academic phrasing, outline structure, paraphrase transitions, and check grammar. After using these tools/services, the author(s) reviewed and edited the content as needed and take(s) full responsibility for the content of the thesis. AI Citations: Google. (2026). \textit{Gemini} [Large language model]. https://gemini.google.com; Anthropic. (2026). \textit{Claude} [Large language model]. https://claude.ai.