% ─────────────────────────────────────────────────────────────────────────────
% Experiment Inventory — two full-width multi-page tables
%
% Required preamble:
%   \usepackage{geometry}    % Required to adjust page margins
%   \usepackage{xltabular}   % Replaces ltablex for robust multi-page tabularx
%   \usepackage{booktabs}
%   \usepackage{array}
% ─────────────────────────────────────────────────────────────────────────────

% Temporarily pull the margins out to 1.5cm for these tables.
% Note: This command automatically triggers a \clearpage
\newgeometry{left=1.5cm, right=1.5cm}

% ── TABLE A: Scientific Purpose ──────────────────────────────────────────────
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.4}

\begin{xltabular}{\textwidth}{
  >{\centering\arraybackslash}p{0.5cm}
  p{2.7cm}
  p{2.0cm}
  >{\centering\arraybackslash}p{1.2cm}
  X
}

\caption{%
  Experiment inventory --- scientific purpose.
  Configs~7 and~8 are already complete.
  $^\dagger$Config~5 is the Sub-RQ3 paired comparison against Config~8;
  the sole difference is the first-task optimizer (Adam vs.\ AdamW).
}\label{tab:exp_purpose} \\

\toprule
\textbf{\#} & \textbf{Benchmark} & \textbf{Variant} &
\textbf{Sub-RQ} & \textbf{What it differentiates} \\
\midrule
\endfirsthead

\multicolumn{5}{c}{\tablename~\thetable{} \textit{--- continued from previous page}} \\[3pt]
\toprule
\textbf{\#} & \textbf{Benchmark} & \textbf{Variant} &
\textbf{Sub-RQ} & \textbf{What it differentiates} \\
\midrule
\endhead

\midrule
\multicolumn{5}{r}{\textit{Continued on next page\ldots}} \\
\endfoot

\bottomrule
\endlastfoot

1 & Permuted-MNIST & Standalone & 4
  & Does iFOPNG improve retention over FOPNG on a domain-incremental
    benchmark with high inter-task Fisher overlap? \\

2 & Split-MNIST (MH) & Standalone & 4, 1a
  & Does iFOPNG replicate and extend FOPNG's canonical result?
    Serves as Panel~(a) unconstrained reference for Sub-RQ1 Benchmark~1. \\

3 & Split-MNIST (SH) & Standalone & 4
  & Does the iFOPNG advantage survive the harder single-head protocol
    \citep{farquhar_towards_2019} where separate task heads do not
    suppress output-level interference? \\

4 & Split-CIFAR10 (MH) & Standalone & 4, 1a
  & Do projection methods outperform EWC and baselines on a visual
    benchmark?
    Serves as Panel~(a) reference for Sub-RQ1 Benchmark~2. \\

5$^\dagger$ & Split-CIFAR10 & HN --- Adam & 3A
  & Does Adam with coupled $L_2$ on the first task contaminate gradient
    memory $G$ relative to AdamW?
    Paired against Config~8 (AdamW first task). \\

6 & Split-CIFAR100 (MH) & Standalone & 4, 1a
  & Does iFOPNG scale to 10 tasks and 10-class per-task splits?
    Serves as Panel~(a) reference for Sub-RQ1 Benchmark~3. \\

\midrule

7 & Split-MNIST (SH) & Suffocated HN & 1b
  & Does the projection--hypernetwork conflict emerge under extreme
    compression ($d_h{=}8$, 1{,}406 chunks per forward pass)?
    Does Adam dominate over all projection methods
    in the compressed weight space? \\

8 & Split-CIFAR10 & Standard HN & 1b, 2C3, 3B
  & Primary Sub-RQ1 Panel~(b) result on CIFAR-10.
    Also Sub-RQ2 Condition~3 (full normalization as proposed) and
    Sub-RQ3 Condition~B (AdamW first task). \\

\midrule

9 & Split-CIFAR10 & HN, no norm & 2C1
  & \textbf{Negative control.}
    Does the chunk-induced gradient pathology produce \texttt{NaN} outputs
    and matrix inversion collapse without any normalization?
    (Expected: yes.) \\

10 & Split-CIFAR10 & HN, grad only & 2C2
   & Is unit-$\ell_2$ gradient normalization alone sufficient to prevent
     inversion collapse, or is Fisher abs-max scaling also required? \\

11 & Split-CIFAR100 & Standard HN & 1b
   & Does the Sub-RQ1 finding (Adam dominates in the hypernetwork
     setting) hold as task count doubles to 10 and per-task
     classification difficulty increases to 10 classes? \\

\midrule

12 & Split-MNIST (SH) & HN $d_h{=}4$ & App.
   & \textit{Prelim.\ sweep lower bound.}
     Does $d_h{=}4$ cause complete training failure across all seeds?
     Justifies the choice of $d_h{=}8$ in Config~7. \\

13 & Split-MNIST (SH) & HN $d_h{=}16$ & App.
   & \textit{Prelim.\ sweep upper bound.}
     Does $d_h{=}16$ produce a ceiling (${\geq}0.93$) with no method
     separation?  Justifies the choice of $d_h{=}8$ in Config~7. \\

\midrule

14 & Permuted-MNIST & EMA Accum. & 5
   & \textbf{Stress Test.} Does Exponential Moving Average (EMA) Fisher 
     accumulation improve retention over MAX accumulation in a long 20-task sequence? \\

15 & Permuted-MNIST & MAX Accum. & 5
   & \textbf{Stress Test.} Paired baseline for Config~14 using standard 
     MAX Fisher construction on a 20-task sequence. \\

16 & Split-MNIST (SH) & HN $\beta$ Sweep & Ablat.
   & \textit{Sweep A.} Determines stable $\beta$ penalty weights for the 
     suffocated hypernetwork regime ($c{=}1000$). \\

17 & Split-CIFAR10 & HN $\beta/c$ Sweep & Ablat.
   & \textit{Sweep B.} Identifies the optimal trade-off between backward transfer 
     (BWT) and accuracy by sweeping $\beta$ and chunk size ($c$). \\

18 & Split-CIFAR100 & HN $\beta$ Sweep & Ablat.
   & \textit{Sweep C.} Evaluates stability under high gradient explosion risk 
     for massive target networks ($c{=}6000$). \\

\midrule

19 & Split-MNIST (SH) & SGD lr sweep & Diag.
   & \textbf{Diagnostic.} At what learning rate does SGD fail to produce a
     usable task-1 initialisation on Split-MNIST within the 5-epoch budget?
     Establishes that the convergence threshold lies between $5{\times}10^{-5}$
     and $10^{-4}$.
     Critically, the paper's own FOPNG learning rate for Split-MNIST is
     $\eta{=}10^{-5}$, which lies a full decade below this threshold
     (SGD at $10^{-5}$ yields acc ${\approx}0.21$, barely above random).
     This directly falsifies the paper's stated first-task protocol
     ``SGD at the same learning rate as the optimizer'' for FOPNG on
     Split-MNIST. \\

20 & Split-CIFAR10 (MH) & SGD lr sweep & Diag.
   & \textbf{Diagnostic.} Identifies the SGD convergence threshold for
     Split-CIFAR10 MH within 5 epochs.
     Result: the paper's stated $\eta{=}10^{-3}$ yields only ${\approx}85\%$
     task-1 accuracy with loss ${\approx}0.65 > \ln 2$, far below the
     ${\approx}95\%$ shown in Figure~1 of \citet{garg_fisher-orthogonal_2026}.
     Threshold is ${\geq}10^{-2}$, directly refuting the paper's claimed
     first-task protocol for CIFAR-10 experiments. \\
     
\end{xltabular}


% ── TABLE B: Execution Details ────────────────────────────────────────────────
\begin{xltabular}{\textwidth}{
  >{\centering\arraybackslash}p{0.5cm}
  p{2.5cm}
  p{2.4cm}
  >{\centering\arraybackslash}p{1.2cm}
  >{\centering\arraybackslash}p{0.7cm}
  X
  >{\centering\arraybackslash}p{1.3cm}
}

\caption{%
  Experiment inventory --- execution details.
  \textbf{Method sets}:
  ALL $=$ \{iFOPNG, FOPNG, OGD, ONG, FNG, EWC, Adam, SGD\};
  PROJ $=$ \{iFOPNG, FOPNG, OGD, ONG, FNG, EWC\};
  FISHER $=$ \{iFOPNG, FOPNG, OGD, FNG\};
  MINI $=$ \{iFOPNG, Adam\}.
  ``Table~1'' $=$ Garg et al.\ (2026) Table~1 exactly (per-method lr and
  $\lambda$). ``HN'' $=$ uniform lr $= 10^{-3}$, $\lambda = 10^{-3}$,
  Fisher $= 1024$, batch $= 64$
  (see Section~\ref{sec:hparam_provenance}).
  $^{*}$EWC in Config~6 uses Fisher $= 5000$ (Table~1 ``full'');
  all other methods use $1024$.
}\label{tab:exp_execution} \\

\toprule
\textbf{\#} & \textbf{Benchmark} & \textbf{First task} &
\textbf{Methods} & \textbf{Seeds} &
\textbf{Notable parameters} & \textbf{Status} \\
\midrule
\endfirsthead

\multicolumn{7}{c}{\tablename~\thetable{} \textit{--- continued from previous page}} \\[3pt]
\toprule
\textbf{\#} & \textbf{Benchmark} & \textbf{First task} &
\textbf{Methods} & \textbf{Seeds} &
\textbf{Notable parameters} & \textbf{Status} \\
\midrule
\endhead

\midrule
\multicolumn{7}{r}{\textit{Continued on next page\ldots}} \\
\endfoot

\bottomrule
\endlastfoot

1 & Permuted-MNIST & SGD @ method lr & ALL & 3
  & Table~1 HPs. Fisher $=$ full (60K). 5 tasks $\times$ 5 ep.
  & \checkmark \\

2 & Split-MNIST (MH) & SGD @ method lr & ALL & 3
  & Table~1 HPs. Fisher $=$ full (12K).
    5 tasks $\times$ 5 ep. Multi-head 5$\times$2.
  & \checkmark \\

3 & Split-MNIST (SH) & SGD @ method lr & ALL & 3
  & Table~1 HPs. Fisher $=$ full (12K).
    5 tasks $\times$ 5 ep. Single 10-class head.
  & \checkmark \\

4 & Split-CIFAR10 (MH) & SGD @ $10^{-3}$ & ALL & 3
  & Table~1 HPs. Fisher $= 1024$.
    5 tasks $\times$ 5 ep. Multi-head 5$\times$2.
  & \checkmark (verify) \\

5 & Split-CIFAR10 HN & Adam @ $10^{-3}$ & FISHER & 3
  & HN (64/32/32/256, 185 chunks). $\beta = 0.01$.
    5 tasks $\times$ 50 ep. Full normalization.
  & \checkmark \\

6 & Split-CIFAR100 (MH) & SGD @ $10^{-2}$ & ALL & 3
  & Table~1 HPs. Fisher $= 1024$ (EWC: 5000$^{*}$).
    max\_dirs $= 800$. 10 tasks $\times$ 10 ep.
  & \checkmark \\

\midrule

7 & Split-MNIST SH HN & AdamW @ $10^{-3}$ & ALL & 5
  & Suffocated HN ($d_h{=}8$, emb $= 32/32$, chunk $= 64$,
    1{,}406 chunks). 5 tasks $\times$ 15 ep. Full normalization.
  & \checkmark \\

8 & Split-CIFAR10 HN & AdamW @ $10^{-3}$ & ALL & 3
  & Standard HN (64/32/32/256, 185 chunks). $\beta = 0.01$.
    5 tasks $\times$ 50 ep. Full normalization.
  & \checkmark \\

\midrule

9 & Split-CIFAR10 HN & AdamW @ $10^{-3}$ & iFOPNG & 3
  & Config~8 settings. \textbf{No} \texttt{--normalize} flag.
    Log cond$(G^\top\hat{F}G)$ per step.
  & It didnt crash for ifopng. The efficiency went down. \\

10 & Split-CIFAR10 HN & AdamW @ $10^{-3}$ & iFOPNG & 3
   & Config~8 settings. \texttt{--normalize\_gradients\_only}
     (unit $\ell_2$ cols; raw Fisher).
   & TODO DEMANDS SMALL SPECIFIC CHANGES. NO FLAG FOR IT. \\

11 & Split-CIFAR100 HN & AdamW @ $10^{-3}$ & ALL & 3
   & Standard HN (64/32/32/256). Same architecture as Config~8.
     10 tasks $\times$ 50 ep. Full normalization.
   & \checkmark (SGD | OGD off. incorporate OGD 1e-4 lr) \\

\midrule

12 & Split-MNIST SH HN & AdamW @ $10^{-3}$ & MINI & 2
   & $d_h{=}4$, emb $= 32/32$, chunk $= 64$.
     5 tasks $\times$ 15 ep.
     Result: Not a total collapse but a degrading performance nonetheless of projection methods.
   & \checkmark \\

13 & Split-MNIST SH HN & AdamW @ $10^{-3}$ & MINI & 2
   & $d_h{=}16$, emb $= 32/32$, chunk $= 64$.
     5 tasks $\times$ 15 ep.
     Result: Both projections and Adam do similarly.
   & \checkmark \\

\midrule

14 & Permuted-MNIST & SGD @ $10^{-3}$ & iFOPNG EMA & 5
   & \textbf{20 tasks} $\times$ 5 ep. Fisher $= 60000$. $\alpha = 0.5$.
     Config 414.
   & \checkmark \\

15 & Permuted-MNIST & SGD @ $10^{-3}$ & iFOPNG & 5
   & \textbf{20 tasks} $\times$ 5 ep. Fisher $= 60000$. $\alpha = 0.5$.
     Config 415.
   & \checkmark \\

\midrule

16 & Split-MNIST SH HN & AdamW @ $10^{-3}$ & Adam & 1
   & \textit{Exp 701.} $c=1000$, $d_h=8$, emb $= 4/10$.
     Sweeps $\beta \in \{0.01, 0.05, 0.1\}$.
   & \checkmark \\

17 & Split-CIFAR10 HN & AdamW @ $10^{-3}$ & Adam & 1
   & \textit{Exp 702.} $d_h=32$, emb $= 16/16$.
     Sweeps $\beta$ and $c \in \{64, 500, 1000, 2000, 5000, 6000\}$.
   & \checkmark \\

18 & Split-CIFAR100 HN & AdamW @ $10^{-3}$ & Adam & 1
   & \textit{Exp 703.} \textbf{Fast check} (5 tasks $\times$ 15 ep).
     $c=6000$, $d_h=128$. Sweeps $\beta$.
   & \checkmark \\
   
\midrule

19 & Split-MNIST (SH) & SGD @ sweep lr & SGD & 5
   & \textbf{Diagnostic.} Exp~419. 1 task $\times$ 5 ep.
     Sweeps $\eta \in \{10^{-5},\allowbreak\ 5{\times}10^{-5},\allowbreak\ 10^{-4},\allowbreak\ 2{\times}10^{-4},\allowbreak\ 5{\times}10^{-4},\allowbreak\ 10^{-3},\allowbreak\ 5{\times}10^{-3},\allowbreak\ 10^{-2}\}$.
     Fisher $= 12000$. Batch $= 10$.
     \textbf{Result:} threshold between $5{\times}10^{-5}$ and $10^{-4}$.
     Paper's FOPNG lr $= 10^{-5}$ $\Rightarrow$ SGD acc ${\approx}0.21$
     (barely above random); protocol is self-contradictory.
   & \checkmark \\

20 & Split-CIFAR10 (MH) & SGD @ sweep lr & SGD & 5
   & \textbf{Diagnostic.} Exp~420. 1 task $\times$ 5 ep.
     Sweeps $\eta \in \{10^{-3},\allowbreak\ 5{\times}10^{-3},\allowbreak\ 10^{-2},\allowbreak\ 5{\times}10^{-2}\}$.
     Fisher $= 1024$. Batch $= 10$.
     \textbf{Result:} paper's $\eta{=}10^{-3}$ yields acc ${\approx}0.85$,
     loss ${\approx}0.65$ (above random $\ln 2$).
     True convergence only at $\eta{\geq}10^{-2}$; refutes paper's
     first-task protocol for CIFAR-10.
   & \checkmark \\
\end{xltabular}

% Return the document to its original margins.
% Note: This command automatically triggers another \clearpage
\restoregeometry