% ─────────────────────────────────────────────────────────────────────────────
% Experimental Design
% Now a \subsection of the Methods section (fits the fixed thesis structure).
%
% Hyperparameter provenance: benchmark-specific values deferred to the
% appendix table.
%
% VERIFY before submission (flagged with % CHECK):
%   - beta value for CIFAR HN (log shows 0.1, old draft said 0.01)
%   - embedding dims (CIFAR HN = 16, MNIST suffocated = 32)
%   - chunk_size / chunk-count for each HN benchmark
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Experimental Design}
\label{sec:experimental_design}

The four sub-research questions demand qualitatively different experimental
designs. Sub-RQ1 and Sub-RQ3 require head-to-head benchmark comparisons ---
projection methods against an unconstrained baseline, and \texttt{iFOPNG}
against \texttt{FOPNG} respectively. Sub-RQ2 requires a controlled ablation
over normalization conditions. Sub-RQ4 requires a long-horizon comparison of
two Fisher accumulation operators on the same elastic preconditioner.
Hyperparameter choices fall into two categories: those inherited directly
from the literature, which require no further justification, and those
specific to the novel hypernetwork experiments, each of which is justified
below.

% ─────────────────────────────────────────────────────────────────────────────
\subsubsection{Hyperparameter Provenance}
\label{sec:hparam_provenance}

\paragraph{Standalone target-network experiments.}
All hyperparameter values for experiments without a hypernetwork ---
learning rates, regularization coefficients, gradient budgets, Fisher
sample counts, batch sizes, and epoch counts --- are taken without
modification from \citet{garg_fisher-orthogonal_2026} Table~1.
\texttt{iFOPNG} uses the same values as \texttt{FOPNG}, since the elastic
combined Fisher introduces no hyperparameters beyond those FOPNG already
defines.
\texttt{ONG} \citep{Yadav_Mendoza_Korrapati_2025} is treated as a composite
method: it applies \texttt{OGD}'s Euclidean projection operator to the
natural gradient direction and introduces no hyperparameters beyond those
already specified for \texttt{OGD} and \texttt{FNG}; its values therefore
follow \texttt{OGD}'s Table~1 entries.
First-task training follows the paper's protocol: SGD at the method's own
learning rate for MNIST-based benchmarks, and SGD at a fixed rate of
$10^{-3}$ for CIFAR-based benchmarks. No tuning was performed on these
values; they are reproduced to enable direct numerical comparison with the
published results.

\paragraph{Hypernetwork experiments.}
The hypernetwork experiments introduce settings not covered by
\citet{garg_fisher-orthogonal_2026}, which did not evaluate projection
methods in a hypernetwork setting.
The functional regularizer and the task and chunk embedding dimensions
follow \citet{oswald_continual_2022}, the canonical hypernetwork continual
learning reference; benchmark-specific values are listed in
Appendix~\ref{app:hyperparameters}.
Fisher estimation is capped at $1{,}024$ samples rather than using the full
training set: a full estimate on a hypernetwork requires one complete
$K$-chunk generation pass per training sample, which is computationally
intractable within the available compute budget. The cap of $1{,}024$ is
consistent with the FOPNG paper's own CIFAR-10 setting and is a standard
practical approximation \citep{kirkpatrick_overcoming_2017}.
The gradient budget is set to $K = 80$ gradients per task, matching the
FOPNG paper's CIFAR values, as no principled reason exists to increase it
for the hypernetwork's similarly-sized target parameter space.
Unlike the standalone experiments, the hypernetwork runs do exceed the
gradient memory budget; compression is handled by the QR-and-SVD memory
management scheme described in Section~\ref{sec:projection_methods}.
First-task training for all hypernetwork experiments uses \texttt{AdamW};
this follows the Sub-RQ1 secondary diagnostic rationale described in
Section~\ref{sec:related_work}, ensuring that gradient memory $G$ and Fisher
estimate $\hat{F}$ are seeded from a geometrically clean first-task basis
rather than one contaminated by coupled $L_2$ weight decay.

\paragraph{Uniform learning rate for hypernetwork experiments.}
Rather than using the method-specific learning rates of Table~1, all methods
in the hypernetwork experiments use a uniform learning rate of $10^{-3}$.
The goal of these experiments is not to replicate Table~1 numbers --- which
apply to standalone networks --- but to isolate the projection constraint
as the sole independent variable. Method-specific rates would conflate
convergence-speed effects with the structural effect of the projection.
The uniform rate $10^{-3}$ was selected as the value at which the
\texttt{Adam} baseline converges reliably within the epoch budget
(task-1 training loss below $0.05$ by epoch~10).

% ─────────────────────────────────────────────────────────────────────────────
\subsubsection{Sub-RQ1 --- Constructive Synergy or Architectural Conflict}
\label{sec:design_rq1}

\paragraph{What we want to differentiate.}
Sub-RQ1 asks whether layering a hard geometric projection on top of a
hypernetwork's soft functional regularizer yields a net benefit, or whether
the momentum-stripping effect of projection methods makes a simple
\texttt{Adam}-regularized hypernetwork a stronger baseline.
To attribute any performance difference to the projection constraint alone
--- rather than to task difficulty, architecture, or training procedure ---
the comparison must hold across a range of difficulty levels and
compression regimes.
A single benchmark cannot answer this convincingly: if it is too easy, all
methods reach a ceiling and no differences are observable; if it is too
hard, the hypernetwork cannot learn and projection methods fail for reasons
unrelated to the research question.
We therefore use a three-benchmark progression, each evaluated under two
panels: Panel~(a), the unconstrained standalone target network, and
Panel~(b), the hypernetwork.

\paragraph{Benchmark 1 --- Split-MNIST with a suffocated hypernetwork.}
A standard hypernetwork on Split-MNIST reaches near-perfect accuracy for all
methods, producing a ceiling effect in which no differential effects are
observable.
To expose differential effects while retaining a benchmark known to be
solvable, the hypernetwork is deliberately \emph{suffocated} by setting the
generator hidden dimension to a minimal value.
The hidden dimension $d_h = 8$ was selected via a preliminary sweep over
$d_h \in \{4, 8, 16, 32\}$ (Appendix~\ref{app:dh_sweep}).
Dimensions $d_h \geq 16$ produced near-ceiling accuracy for all methods
($>0.93$), leaving no measurable separation; $d_h = 4$ caused complete
training failure across all methods and seeds (random-chance accuracy
$\approx 0.10$).
$d_h = 8$ was the only configuration producing measurable separation
between methods while remaining above the failure threshold for the
majority of seeds.
The chunk size of $64$ produces $1{,}406$ weight-generation passes per
forward call (target network $\approx 90{,}000$ parameters), maximally
stressing the chunk-induced gradient accumulation pathology so that any
difference between normalized and unnormalized projection is also visible
in this panel.
Each task is trained for $15$ epochs: preliminary runs showed the
suffocated hypernetwork does not converge within the FOPNG paper's
$5$-epoch budget (task-1 loss remained above $0.20$ at epoch~5 and continued
decreasing through epoch~12). Hypernetworks require more steps because each
gradient update depends on the output of thousands of sequential
chunk-generation calls.
The batch size is set to $64$ rather than the canonical $10$: each
hypernetwork training step regenerates all $1{,}406$ chunks before a single
gradient is computed, and the larger batch reduces the step count
proportionally while fitting the GPU memory budget; the method ordering is
preserved regardless of batch size.
Panel~(a) uses the multi-head MLP standalone target with method-specific
Table~1 learning rates, establishing the unconstrained reference and
validating implementation correctness.

\paragraph{Benchmark 2 --- Split-CIFAR10 with a standard hypernetwork.}
Split-CIFAR10 is the primary benchmark for Sub-RQ1. Its higher input
complexity creates genuine inter-task interference, preventing ceiling
effects without reaching the regime where the hypernetwork fails entirely.
The hypernetwork configuration --- generator hidden dimension, chunk size,
and embedding dimensions --- is the smallest setting for which the
hypernetwork achieves $\geq 70\%$ accuracy on task~1 with \texttt{Adam},
establishing sufficient capacity to learn before the projection constraint
is applied; exact values are listed in Appendix~\ref{app:hyperparameters}.
Each task is trained for $50$ epochs, consistent with \citet{oswald_continual_2022}'s
recommendation of extended training for CIFAR-scale hypernetworks;
preliminary runs confirmed $5$-epoch convergence was insufficient.
Panel~(a) uses the FOPNG-canonical multi-head CNN with Table~1
hyperparameters.

\paragraph{Benchmark 3 --- Split-CIFAR100 with a standard hypernetwork.}
Split-CIFAR100 extends the task sequence to $10$ tasks and the per-task
classification problem to $10$ classes, testing whether the Sub-RQ1 finding
scales with sequence length and task complexity.
The hypernetwork configuration matches Benchmark~2.
Panel~(a) uses the same CNN backbone with ten $10$-class output heads.
If the \texttt{Adam}-versus-projection gap grows with task count, this
indicates that momentum becomes increasingly critical as the hypernetwork's
task manifold grows more complex.

\paragraph{Three outcome patterns.}
Comparing Panel~(a) against Panel~(b) within each benchmark, and across
benchmarks, distinguishes three outcomes:
\begin{enumerate}
    \item \emph{Projection methods dominate in both panels.}
    The combination is constructive; Sub-RQ1 is answered affirmatively.
    \item \emph{Projection methods dominate in Panel~(a) but \texttt{Adam}
    dominates in Panel~(b).}
    The momentum-stripping incompatibility is the primary mechanism;
    Sub-RQ1 is answered negatively.
    \item \emph{Results depend on compression regime or task difficulty.}
    The integration is conditionally constructive, warranting finer-grained
    analysis of when projection adds value.
\end{enumerate}

% ─────────────────────────────────────────────────────────────────────────────
\subsubsection{Sub-RQ2 --- Projection-Scoped Normalization}
\label{sec:design_rq2}

\paragraph{What we want to differentiate.}
Sub-RQ2 asks whether the chunk-induced gradient accumulation pathology is
real, characterisable, and resolvable.
A single accuracy number cannot answer this: we need direct numerical
evidence that the pathology occurs without normalization, that
normalization resolves it, and that the resolution does not compromise
projection quality.

\paragraph{Experimental design.}
\texttt{iFOPNG} is run on the Split-CIFAR10 hypernetwork (Benchmark~2)
under three normalization conditions, all other hyperparameters held
constant:
\begin{description}
    \item[Condition 1 --- No normalization (negative control).]
    Incoming gradient batches are appended to $G$ without orthonormalization;
    the Fisher diagonal is used at its raw scale.
    With several hundred chunks accumulated per forward pass, the gradient
    columns of $G$ and the entries of $\hat{F}$ are expected to reach
    extreme magnitudes, inflating the condition number of the projection
    kernel and driving the matrix inversion to \texttt{NaN}. This condition
    is the negative control that demonstrates the pathology concretely.
    \item[Condition 2 --- Gradient orthonormalization only.]
    Each incoming gradient batch is orthonormalized via QR decomposition
    before insertion into $G$; the Fisher diagonal is used at its raw scale.
    This isolates whether gradient-side conditioning alone is sufficient, or
    whether Fisher-side scaling is also required.
    \item[Condition 3 --- Full normalization (proposed).]
    Gradient batches are QR-orthonormalized at insertion, and the Fisher
    terms are scaled by the maximum entry of the metric tensor $F_c$ within
    each projection computation. This is the proposed scheme; all other
    hypernetwork experiments use it.
\end{description}
In addition to accuracy and BWT, the condition number of the projection
kernel $A$ is logged before each inversion.
A condition number that diverges in Condition~1 but remains bounded in
Condition~3 constitutes direct numerical evidence for the pathology and its
resolution.
All conditions use the Benchmark~2 hypernetwork settings and the uniform
learning rate of $10^{-3}$, so any performance difference is attributable
to normalization alone.

% ─────────────────────────────────────────────────────────────────────────────
\subsubsection{Sub-RQ3 --- Does iFOPNG's Elastic Preconditioner Help}
\label{sec:design_rq3}

\paragraph{What we want to differentiate.}
Sub-RQ3 asks whether the elastic combined Fisher
$F_c = \hat{F}_{\mathrm{new}} + \hat{F}_{\mathrm{old}}$ provides a
consistent retention improvement over rigid FOPNG, and whether any
advantage is robust to the evaluation protocol.

\paragraph{Benchmark selection.}
The full method suite is evaluated across the standalone benchmarks, each
chosen to isolate a specific dimension of the comparison:
\begin{description}
    \item[Permuted-MNIST.] Domain-incremental, single shared head, high
    inter-task Fisher overlap. Tests long-horizon retention under
    input-distribution shift. If $F_c$ accumulates historical curvature
    effectively, \texttt{iFOPNG} should show smaller BWT than \texttt{FOPNG}
    precisely where Fisher overlap is high.
    \item[Split-MNIST (multi-head, \citealt{garg_fisher-orthogonal_2026}).]
    Canonical multi-head evaluation with five $2$-class heads and remapped
    labels, replicating the conditions under which FOPNG reported its own
    results and providing a direct numerical comparison point.
    \item[Split-MNIST (single-head, \citealt{farquhar_towards_2019}).]
    The same five binary tasks presented to a single shared $10$-class head
    with original labels preserved. \citet{farquhar_towards_2019} showed
    that multi-head evaluation inflates apparent performance by removing
    output-level interference; this variant tests whether the
    \texttt{iFOPNG} advantage survives the harder protocol.
    \item[Split-CIFAR10 and Split-CIFAR100 (multi-head).]
    The most discriminative standalone benchmarks: stronger inter-task
    interference than MNIST, clear method separation, and direct
    comparability with the Panel~(a) references of Sub-RQ1.
    Split-CIFAR100 additionally tests whether the elastic advantage holds
    as the sequence extends to $10$ tasks.
\end{description}

\paragraph{What a difference reveals.}
Comparing \texttt{iFOPNG} against \texttt{FOPNG} across all benchmarks tests
whether the elastic preconditioner provides a consistent, robust improvement
or a benchmark-specific artefact. Comparing both against \texttt{OGD} tests
whether the Fisher metric contributes any benefit beyond Euclidean
projection. Comparing the two Split-MNIST variants tests the
protocol-robustness claim of \citet{farquhar_towards_2019} directly.
All hyperparameters for these experiments follow
\citet{garg_fisher-orthogonal_2026} Table~1 as described in
Section~\ref{sec:hparam_provenance}; no tuning was performed.

% ─────────────────────────────────────────────────────────────────────────────

\subsubsection{Sub-RQ4 --- Fisher Accumulation Operator}
\label{sec:design_rq4}

\paragraph{What we want to differentiate.}
Sub-RQ4 holds the elastic formulation of Sub-RQ3 fixed and isolates the
operator used to accumulate the historical Fisher $F_{\mathrm{old}}$.
The non-decaying maximum $F_{\mathrm{old}} \leftarrow
\max(F_{\mathrm{old}}, F_{\mathrm{new}})$ guarantees that any parameter
ever deemed important to a past task retains that importance permanently.
This is protective for early tasks but monotonic by construction: the
elastic inertia boundary can only tighten, never relax, so in a
sufficiently long sequence the accumulated metric may immobilise the
parameters needed to learn later tasks.

\paragraph{Experimental design.}
\texttt{iFOPNG} is run under two accumulation operators, all other settings
held constant:
\begin{description}
    \item[Maximum (proposed).] $F_{\mathrm{old}} \leftarrow
    \max(F_{\mathrm{old}}, F_{\mathrm{new}})$ --- non-decaying.
    \item[Exponential moving average.] $F_{\mathrm{old}} \leftarrow
    (1-\alpha)\,F_{\mathrm{old}} + \alpha\,F_{\mathrm{new}}$ ---
    decaying, allowing historical importance to fade.
\end{description}
The comparison is run on the longest available task sequences ---
Permuted-MNIST and Split-CIFAR100 --- since the plasticity cost of
monotonic accumulation, if present, only becomes observable as the
sequence length grows.
Per-task accuracy is tracked across the full sequence rather than reported
only as a final aggregate: the diagnostic signature of plasticity paralysis
is a progressive decline in \emph{newly learned} task accuracy
($a_{t,t}$ falling as $t$ increases), distinct from forgetting, which
appears as decline in \emph{previously learned} task accuracy.
If the maximum operator shows superior early-task retention but degrading
$a_{t,t}$ in late tasks, while the moving average shows the opposite trade,
the two failure modes are cleanly separated and Sub-RQ4 is answered in
full.