% ─────────────────────────────────────────────────────────────────────────────
% METHODOLOGY SECTION — clean rewrite
% Principles:
%   - Operational focus: equations + implementation choices, not re-introduction
%   - Related Work already introduced EWC, OGD, FOPNG, HN conceptually
%   - Novel contributions (iFOPNG, normalization, dropout decision) get full
%     treatment; baselines get equations + key parameters only
%   - ONG and FNG filled in from Garg et al. (2026) formulations
%   - Dropout paragraph cleaned of HTML rendering artifacts
% ─────────────────────────────────────────────────────────────────────────────

\section{Methods}
\label{sec:methods}

This section specifies the precise formulation of every method and
architectural component used in our experiments, providing sufficient
detail for reproducibility.
Conceptual motivation and positioning relative to prior work are covered
in Section~\ref{sec:related_work}; we restrict our attention here to
the exact algorithmic instantiations evaluated.

% ─────────────────────────────────────────────────────────────────────────────
% ─────────────────────────────────────────────────────────────────────────────
\subsection{Continual Learning Protocol}
\label{sec:cl_protocol}

\subsubsection{Task Setup and Evaluation Metrics}
A continual learning agent receives a sequence of $T$ tasks
$\mathcal{T}_1, \mathcal{T}_2, \ldots, \mathcal{T}_T$ presented in order.
At task $t$, only the data from $\mathcal{T}_t$ is available; prior task
data is not retained.
After each task is trained, the model is evaluated on all tasks seen so far.
We report two aggregate metrics: \emph{average accuracy}
\begin{equation}
    A_t = \frac{1}{t}\sum_{i=1}^{t} a_{t,i},
\end{equation}
and \emph{backward transfer}
\begin{equation}
    \mathrm{BWT} = \frac{1}{T-1}\sum_{t=2}^{T}(a_{T,t}-a_{t,t}),
\end{equation}
where $a_{i,j}$ denotes accuracy on task $j$ immediately after training
task $i$.
Negative BWT quantifies catastrophic forgetting; a value of zero indicates
perfect retention.

\subsubsection{Multi-Head and Single-Head Evaluation}
Following \citet{farquhar_towards_2019}, who demonstrated that multi-head
evaluations can inflate apparent method performance, we report results under
both protocols for MNIST-based benchmarks.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Fisher Information Matrix and Regularization-Based Continual Learning}
\label{sec:fisher_ewc}

\subsubsection{Fisher Information Matrix}
All Fisher-based methods in this thesis use the \emph{diagonal empirical
approximation}
\begin{equation}
    \hat{F}_i \approx \frac{1}{N}\sum_{n=1}^{N}
    \!\left(\frac{\partial \mathcal{L}^{(n)}}{\partial \theta_i}\right)^{\!2},
\end{equation}
estimated over $N$ held-out samples at the end of each task.
The full FIM is intractable for our network sizes; the diagonal
approximation is standard across all evaluated methods
\citep{kirkpatrick_overcoming_2017,garg_fisher-orthogonal_2026}.

\subsubsection{Elastic Weight Consolidation}
\label{sec:ewc}
EWC \citep{kirkpatrick_overcoming_2017} augments the task loss with a
Fisher-weighted quadratic penalty:
\begin{equation}
    \mathcal{L}_{\mathrm{EWC}}(\theta) = \mathcal{L}_t(\theta)
    + \frac{\lambda}{2}\sum_i \hat{F}_i\,(\theta_i - \theta_{i,t-1}^*)^2,
\end{equation}
where $\theta^*_{t-1}$ denotes the parameter snapshot after task $t{-}1$.
EWC operates exclusively through parameter-space penalties and maintains
no gradient subspace; it is included as a regularization baseline employing
a different mechanism from projection.
We use $\lambda = 10^{-3}$ for all hypernetwork configurations and
reproduce the per-benchmark values from \citet{garg_fisher-orthogonal_2026}
Table~1 for standalone evaluations.
Fisher estimates are computed over the full training set for MNIST-based
benchmarks and over $N{=}1024$ samples for CIFAR benchmarks.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Gradient Projection Methods}
\label{sec:projection_methods}

This subsection presents the full suite of gradient-based methods evaluated
in this thesis.
\texttt{SGD} and \texttt{FNG} establish unconstrained lower bounds with no
forgetting-protection mechanism; \texttt{OGD} through \texttt{iFOPNG} form
the projection family that is the primary focus of this work.

\subsubsection{Stochastic Gradient Descent}
\texttt{SGD} applies vanilla gradient descent with no forgetting protection:
\begin{equation}
    \Delta\theta_{\mathrm{SGD}} = -\eta\,\nabla_\theta\mathcal{L}_t.
\end{equation}
\texttt{Adam} \citep{kingma_adam_2014} is included as an adaptive-rate
variant.
Both define the lower bound on retention performance; any continual learning
method should improve upon \texttt{SGD} on the forgetting axis.

\subsubsection{Natural Gradient and Fisher Natural Gradient}
Natural gradient descent \citep{amari_natural_1998} preconditions the
gradient with the inverse FIM, respecting the curvature of the statistical
manifold:
\begin{equation}
    \Delta\theta_{\mathrm{nat}} = -\eta\,\hat{F}^{-1}\nabla_\theta\mathcal{L}.
\end{equation}
With a diagonal Fisher approximation, inversion reduces to element-wise
scaling and remains computationally tractable.

\texttt{FNG} \citep{garg_fisher-orthogonal_2026} applies a trust-region
constraint to this Riemannian step:
\begin{equation}
    v^*_{\mathrm{FNG}}
    = -\eta\,\frac{\hat{F}^{-1}g}{\sqrt{g^\top \hat{F}^{-1} g}}.
\end{equation}
FNG imposes no projection subspace and constitutes no continual learning
mechanism; it is included to isolate the contribution of Fisher
preconditioning from that of gradient projection.

\subsubsection{Orthogonal Gradient Descent}
\label{sec:ogd}
OGD \citep{farajtabar_orthogonal_2019} maintains a memory matrix
$G = [g_1 \mid \cdots \mid g_K] \in \mathbb{R}^{p \times K}$ of past task
gradient directions and updates parameters via the current gradient
projected onto the orthogonal complement of $\operatorname{Col}(G)$:
\begin{equation}
    \tilde{g} = g - G(G^\top G)^{-1}G^\top g,
    \qquad
    \Delta\theta_{\mathrm{OGD}} = -\eta\,\tilde{g}.
\end{equation}
OGD operates in Euclidean space, treating all parameter directions
symmetrically.

\paragraph{Gradient memory management.}
In all projection methods, $G$ is maintained as a fixed-budget basis of at
most $M$ columns.
\citet{garg_fisher-orthogonal_2026} set $M = T \times K$ across all
benchmarks, so compression never triggers in their experiments.
We address this unexamined regime explicitly: incoming gradient batches are
appended directly to $G$; when the column count exceeds $M$, the full basis
is recompressed via truncated SVD and $G \leftarrow U_{:,\,1:M}$, retaining
the $M$ directions of greatest variance in the accumulated gradient memory.
When projection-scoped normalization is enabled, QR decomposition is
additionally applied to each incoming batch at insertion time
(Section~\ref{sec:normalization}).

\subsubsection{Orthogonal Natural Gradient}
ONG applies OGD-style projection to the natural gradient rather than to the
raw gradient, combining Fisher preconditioning with a Euclidean orthogonality
constraint:
\begin{equation}
    \Delta\theta_{\mathrm{ONG}}
    = -\eta\!\left(
        \hat{F}^{-1}g - G(G^\top G)^{-1}G^\top\hat{F}^{-1}g
      \right).
\end{equation}
Unlike FOPNG, the projection is measured in Euclidean rather than Fisher
metric space, so the orthogonality guarantee does not extend to the
statistical manifold.
ONG uses the same memory budget as OGD.

\subsubsection{Fisher-Orthogonal Projected Natural Gradient Descent}
\texttt{FOPNG} \citep{garg_fisher-orthogonal_2026} executes projection in
the Riemannian metric defined by the Fisher:
\begin{equation}
    v^*_{\mathrm{FOPNG}} = -\eta\,
    \frac{\hat{F}^{-1}Pg}{\sqrt{(Pg)^\top\hat{F}^{-1}Pg}},
    \qquad
    P = I - \hat{F}G\!\left(G^\top\hat{F}G + \lambda I\right)^{\!-1}G^\top,
    \label{eq:fopng}
\end{equation}
where $\lambda$ is added to the diagonal for numerical stability.
We reproduce the per-benchmark learning rates and $\lambda$ values from
\citet{garg_fisher-orthogonal_2026} Table~1 for standalone configurations,
and use a uniform $\mathrm{lr}{=}10^{-3}$, $\lambda{=}10^{-3}$ for all
hypernetwork configurations.

\subsubsection{Inertial FOPNG}
\label{sec:ifopng}

In FOPNG, the parameter update takes the form
\begin{equation}
    v = c\,\hat{F}_{\mathrm{new}}^{-1}(g - u),
    \label{eq:fopng_update_form}
\end{equation}
where $\hat{F}_{\mathrm{new}}$ serves as the Riemannian metric: both the
trust-region constraint $\|v\|_{\hat{F}_{\mathrm{new}}} = \eta$ and the
projection objective $\|g - \hat{F}_{\mathrm{new}}^{-1}u\|^2_{\hat{F}_{\mathrm{new}}}$
are measured in the current task's Fisher geometry.
This is natural when no prior tasks exist, but creates an asymmetry once
history accumulates: the constraint subspace $\mathrm{Col}(G)$ is built from
gradient directions collected under $\hat{F}_{\mathrm{old}}$, yet the
projection metric assigns zero curvature to parameters that were important
in previous tasks but uninformative on the current one.

\texttt{iFOPNG} is introduced in this thesis as a variant of FOPNG that
resolves this asymmetry by adopting the \emph{combined metric}
\begin{equation}
    F_c \;\triangleq\; \hat{F}_{\mathrm{new}} + \hat{F}_{\mathrm{old}},
    \label{eq:combined_fisher}
\end{equation}
where $\hat{F}_{\mathrm{old}}$ is maintained as the element-wise maximum of
all per-task Fisher diagonals accumulated so far.
Replacing \eqref{eq:fopng_update_form} with $v = c\,F_c^{-1}(g - u)$, the
constrained trust-region update becomes
\begin{equation}
    v^* = -\eta\,
    \frac{F_c^{-1}Pg}{\sqrt{(Pg)^\top F_c^{-1}Pg}},
    \qquad
    P = I - \hat{F}_{\mathrm{old}}\,G\,A^{-1}\,G^\top\hat{F}_{\mathrm{old}},
    \label{eq:ifopng}
\end{equation}
where $A = G^\top\hat{F}_{\mathrm{old}}\,F_c^{-1}\,\hat{F}_{\mathrm{old}}\,G$.
A formal derivation is provided in Appendix~A.

The combined metric $F_c$ assigns each coordinate $i$ an effective mass
$(\hat{F}_{\mathrm{new}})_i + (\hat{F}_{\mathrm{old}})_i$, suppressing the
natural-gradient step size in directions important to any prior task.
We call this property \emph{parameter inertia}: accumulated Fisher importance
makes a coordinate harder to displace without any explicit penalty term.
In direct contrast to EWC \citep{kirkpatrick_overcoming_2017}, which pulls
parameters toward a stored reference point $\theta^*$, \texttt{iFOPNG}
maintains no reference point: a parameter that mattered before is not pulled
back, it is simply made harder to displace.
% 
% ─────────────────────────────────────────────────────────────────────────────
\subsection{Hypernetwork Architecture}
\label{sec:hn_arch}

\subsubsection{Chunked Hypernetworks}
A hypernetwork \citep{ha_hypernetworks_2016} is a meta-model
$\mathcal{H}(\cdot;\phi)$ that generates the complete weight vector
$\theta_t = \mathcal{H}(e_t;\phi)$ of a target network conditioned on
a task embedding $e_t$.
Generating $\theta_t$ in a single forward pass would require an output
layer of dimension $|\theta_t|$, which is computationally prohibitive for
any non-trivial target architecture.
\emph{Weight chunking} \citep{oswald_continual_2022} resolves this by
iterating over a secondary chunk embedding $c_k$, $k=1,\ldots,K$,
and producing a block of $C_s$ target parameters per pass; the full weight
vector is assembled by concatenation.
Chunking makes the hypernetwork parameter-efficient but introduces the
chunk-induced gradient accumulation pathology addressed in
Section~\ref{sec:normalization}.

\subsubsection{Functional Regularizer}
The functional regularizer of \citet{oswald_continual_2022} is active for
all hypernetwork configurations, with regularization strength $\beta$ treated
as a tunable hyperparameter (see Appendix~\ref{app:hyperparameters}).
\subsubsection{Target Network and Dropout}
\label{sec:dropout}
Standalone configurations reproduce the target network architecture of
\citet{garg_fisher-orthogonal_2026} without modification, including two
\texttt{Dropout}$(p{=}0.5)$ layers in the \texttt{SimpleCIFARCNN} classifier
head.
Dropout is removed from the target network in all hypernetwork configurations:
\texttt{PyTorch}'s \texttt{functional\_call} propagates the parent module's
\texttt{.training} mode to the target submodule, causing stochastic masking
across all chunk-generation passes during training.
The generative bottleneck $(\phi: d_t{+}d_c \to d_h \to C_s)$ and the
functional regularizer $\mathcal{L}_\mathrm{reg}$ provide sufficient implicit
regularization in its place.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Novel Algorithmic Contributions}
\label{sec:novel}

\subsubsection{Projection-Scoped Normalization}
\label{sec:normalization}

When a hypernetwork spawns $K$ sequential weight-chunks before a single
projection step executes, the accumulated gradient vector and the diagonal
Fisher entries grow in magnitude proportionally to $K$, inflating the
condition number of the projection kernel $G^\top\hat{F}G$ and causing the
matrix inversion in Equation~\eqref{eq:fopng} to fail catastrophically for
any practically sized target network.

We resolve this with two independent rescaling operations that address
distinct sources of ill-conditioning:

\begin{enumerate}
    \item \textbf{Gradient basis orthonormalization.}
        Each batch of $K$ gradient vectors collected after task $t$ is
        replaced by the orthonormal factor $Q$ from its thin QR
        decomposition before insertion into $G$:
        \begin{equation}
            [g_{t,1} \mid \cdots \mid g_{t,K}]
            \;\xrightarrow{\;\mathrm{QR}\;}\; Q,
            \qquad Q^\top Q = I.
        \end{equation}
        QR is chosen over SVD at insertion time because the operation
        requires only orthonormalization of the incoming batch --- the
        number of columns is preserved, not reduced, so the ordered
        factorization that SVD provides is unnecessary overhead.
        SVD is reserved for the overflow handler: when the column count
        of $G$ exceeds the budget $M$, the full basis is recompressed as
        $G \leftarrow U_{:,\,1:M}$, retaining the $M$ left singular
        vectors of greatest variance and discarding the rest.
        This operation applies \emph{at storage time} and affects only
        the basis $G$; the gradient $g$ entering the projection update
        itself is the unmodified parameter gradient from the model.

    \item \textbf{Metric-tensor normalization.}
          Within each projection computation, both Fisher terms are
          divided by the maximum entry of the metric tensor $F_c$ that
          defines the constrained natural gradient problem:
          \begin{equation}
              \tilde{F} = \frac{F}{\max_i (F_c)_i + \varepsilon},
              \qquad F_c = \hat{F}_{\mathrm{new}} + \hat{F}_{\mathrm{old}},
          \end{equation}
          where $F_c$ is the combined Fisher used as the Riemannian metric
          in iFOPNG, or $\hat{F}_{\mathrm{new}}$ alone for FOPNG.
          This scaling is applied locally within the kernel computation
          and does not affect the optimisation trajectory outside the
          projection step.
\end{enumerate}

Together, these operations ensure $G^\top\tilde{F}G$ remains
well-conditioned regardless of the number of chunks spawned per task.
Gradient orthonormalization eliminates column-scale inflation in $G$;
metric-tensor normalization bounds the Fisher entries to $[0, 1]$,
preventing the kernel from becoming numerically singular.
Subspace membership is invariant to positive scalar rescaling of basis
vectors, so neither operation alters the geometry of the projection.


% \paragraph{Principled rescaling and open problem.}
% \label{sec:ng_limitation}
% The proposed normalization — QR orthonormalization of the gradient basis
% combined with max-entry scaling of $F_c$ — is heuristic in nature: it
% discards the eigenvalue structure of the pullback metric and applies
% isotropic rather than curvature-aware rescaling to the chunk embedding
% directions.
% The principled alternative is the natural gradient under the Riemannian
% metric induced by the chunk-generating map:
% \begin{equation}
%     \Delta c_i = -\eta\,(J_i^\top J_i + \lambda I)^{-1}\,
%                   \nabla_{c_i}\mathcal{L},
%     \qquad
%     J_i = \frac{\partial W_i}{\partial c_i} \in \mathbb{R}^{C_s \times d_c},
% \end{equation}
% which rescales each chunk embedding direction in proportion to how
% strongly it moves the generated weights, and is tractable because
% $d_c \ll C_s$.
% This preconditioner was implemented and evaluated experimentally.
% Applied to the raw gradient \emph{prior} to the projection step, it
% consistently degraded continual learning performance — BWT worsened by
% approximately $0.33$ on Split-MNIST SH relative to the heuristic baseline.
% We attribute this to a metric mismatch: the FOPNG projection operator is
% calibrated for gradients in Euclidean parameter space, and preconditioning
% only the chunk embedding component before projection produces a
% geometrically inconsistent input to the kernel $G^\top\hat{F}G$.
% The principled resolution — applying $(J_i^\top J_i + \lambda I)^{-1}$
% to the projected update $v^*$ rather than to the raw gradient, so that the
% projection sees Euclidean gradients and the Riemannian step follows — is
% left to future work.