% ─────────────────────────────────────────────────────────────────────────────
% Related Work — rewrite
%
% Narrative flow (user-specified):
%   1. CL problem definition, why it matters, taxonomy (foundational papers)
%   2. Methods for task-incremental CL
%   3. Multi-head / single-head evaluation design
%   4. EWC as gold standard → projection methods as harder alternative
%   5. Projection methods in detail
%   6. Hypernetworks replacing multi-head design
%   7. Oswald: EWC destructive for HNs (key motivating finding)
%   8. Primary research gap: projection methods + HNs
%   9. iFOPNG as additional contribution
%
% Citation keys used match those in the existing .tex files.
% Sub-RQ5 added to the Identified Gaps section.
% ─────────────────────────────────────────────────────────────────────────────

\section{Related Work}
\label{sec:related_work}

This section builds the theoretical and empirical context for the thesis,
tracing a path from foundational continual learning challenges through
regularization and projection methods, to hypernetwork architectures and the
structural gap at their intersection.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Continual Learning: Taxonomy and the Stability-Plasticity Dilemma}

Modern machine learning systems are predominantly trained on static, i.i.d.\ 
datasets --- a condition that rarely holds in deployment, where agents must
adapt to non-stationary data streams while retaining prior knowledge.
The fundamental tension between these two objectives is formalized as the
\emph{stability-plasticity dilemma} \citep{grossberg_how_1980}: an agent must
remain plastic enough to absorb novel distributions yet stable enough to resist
the erasure of established representations.
When standard architectures trained by empirical risk minimization are exposed
to sequential data, they suffer \emph{catastrophic forgetting}
\citep{mccloskey_catastrophic_1989} --- abrupt and near-total loss of
previously learned capabilities upon exposure to new task data.

The continual learning literature addresses this across three canonical
scenarios \citep{de_lange_continual_2022}:
\emph{task-incremental learning}, where the task identity is available at
both training and test time; \emph{class-incremental learning}, where task
identity is withheld at inference; and \emph{domain-incremental learning},
where the output space is shared but the input distribution shifts
continuously.
This thesis focuses on the task-incremental setting, the scenario in which
projection-based methods and hypernetwork architectures have been most
thoroughly evaluated and in which structural comparisons between them are
most tractable.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Evaluation Protocols: Multi-Head and Single-Head Designs}

Within the task-incremental setting, two evaluation protocols govern how the
network's output layer is organized.
In the \emph{multi-head} protocol, each task is assigned a dedicated output
head; at test time the correct head is selected using the known task identity.
This eliminates output-level interference between tasks but requires that task
identity be available at inference time, a condition that may be unrealistic
in deployment.
In the \emph{single-head} protocol, all tasks share one output layer,
requiring the backbone alone to resolve inter-task interference --- a
substantially harder problem.

\citet{farquhar_towards_2019} demonstrated that multi-head evaluations
systematically inflate apparent method performance by removing a major source
of forgetting, and argued that single-head evaluation is the more rigorous
protocol for assessing continual learning efficacy.
This distinction has direct implications for how gradient projection methods
and hypernetworks are evaluated: projection methods were predominantly
developed and benchmarked in multi-head settings, while hypernetworks
implement an implicit form of task separation through their generative
structure.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Regularization-Based Approaches}

The dominant paradigm for addressing catastrophic forgetting is
\emph{parameter regularization}, of which Elastic Weight Consolidation (EWC)
\citep{kirkpatrick_overcoming_2017} is the canonical representative.
EWC estimates the structural importance of each weight using the diagonal
of the empirical Fisher Information Matrix $\hat{F}$, and applies a soft
quadratic penalty that resists changes to parameters important to prior tasks:
\begin{equation}
    \mathcal{L}_{\mathrm{EWC}}(\theta)
    = \mathcal{L}_t(\theta)
    + \frac{\lambda}{2}\sum_i \hat{F}_i
      \bigl(\theta_i - \theta_{i,t-1}^*\bigr)^2.
\end{equation}
EWC is computationally efficient and broadly applicable, and serves as the
standard reference point against which new continual learning methods are
measured.
Its principal limitation is that soft penalties permit bounded parameter
drift: cumulative updates across a long task sequence gradually erode the
protection of earlier task spaces, a failure mode that becomes increasingly
pronounced as $T$ grows \citep{farquhar_towards_2019}.

A related limitation concerns baseline comparisons.
The superiority of complex regularization schemes over vanilla optimizers
is frequently overstated when those baselines are under-tuned.
When SGD and Adam baselines are tuned to their optimal hyperparameter regime,
they often present a competitive threshold that challenges
or matches more sophisticated approaches --- an empirical reality that our
baseline evaluations directly address.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Gradient Projection Methods}

To enforce strict, non-negotiable protection for prior knowledge rather than
relying on penalties that can be overcome by sufficiently large gradients,
a second paradigm uses explicit \emph{gradient projection}.
Orthogonal Gradient Descent (OGD) \citep{farajtabar_orthogonal_2019} maintains
a memory matrix $G = [g_1 \mid \cdots \mid g_K]$ of historical task gradient
directions and projects incoming gradients onto the orthogonal complement of
$\operatorname{Col}(G)$ before each update:
\begin{equation}
    \tilde{g} = g - G(G^\top G)^{-1}G^\top g.
\end{equation}
This provides a hard guarantee: the update cannot increase the loss on
previously seen tasks.
Generalization bounds established within the Neural Tangent Kernel regime
confirm that OGD provides exact protection in linear and mildly nonlinear
settings \citep{bennani_generalisation_2020}.

The mathematical appeal of projection methods over EWC lies precisely in this
hardness: where EWC applies a soft penalty that can be overwhelmed, projection
methods impose a geometric constraint that is either satisfied or not.
However, standard Euclidean projection treats the parameter space as flat and
assigns equal importance to all directions.
Fisher-Orthogonal Projected Natural Gradient Descent (FOPNG)
\citep{garg_fisher-orthogonal_2026} addresses this by performing the
projection in the Riemannian geometry defined by the Fisher metric, so that
the orthogonality constraint respects the curvature of the loss surface.
The related Fisher Natural Gradient (FNG) \citep{garg_fisher-orthogonal_2026}
applies the natural gradient without projection, and Orthogonal Natural
Gradient (ONG) \citep{Yadav_Mendoza_Korrapati_2025} combines natural gradient
preconditioning with Euclidean projection --- sitting structurally between OGD
and FOPNG.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Hypernetworks for Continual Learning}
An orthogonal architectural response to catastrophic forgetting replaces the
multi-head output design entirely with a \emph{hypernetwork}
\citep{ha_hypernetworks_2016}: a meta-model
$\mathcal{H}(\cdot;\Theta_h)$ trained to synthesize the complete weight
vector $\theta_t = \mathcal{H}(e_t;\Theta_h)$ of a target network,
conditioned on a task embedding $e_t$.
Each task receives a unique weight configuration generated on demand, rather
than a dedicated output head, achieving implicit task separation through the
generative bottleneck and impressive parameter compression ratios relative to
storing separate networks per task.
Generating the full $\theta_t$ in a single forward pass is computationally
intractable for large target networks, so \citet{oswald_continual_2022}
introduced \emph{weight chunking}: the generator iterates over a secondary
chunk embedding $c_k$, producing $C_s$ target parameters per pass and
assembling the full weight vector by concatenation.
To prevent the hypernetwork from forgetting how to generate prior task
configurations, a soft \emph{functional regularizer} is applied in the output
space of the generator:
\begin{equation}
    \mathcal{L}_{\mathrm{reg}}
    = \beta \sum_{s=1}^{t-1}
      \bigl\|\mathcal{H}(\Theta_h,e_s) - \theta_s^*\bigr\|_2^2,
\end{equation}
where $\theta_s^*$ denotes the weight snapshot saved after task $s$.
This penalty can be interpreted as EWC applied at the meta level: rather than
anchoring individual network parameters, it anchors the generated weight
configurations for each previously seen task, penalizing drift in the
space where task-relevant quantities actually live.
Recent extensions have refined the hypernetwork approach: Partial
Hypernetworks \citep{hemati_partial_2023} reduce the computational overhead of
full weight generation, and Prototype Augmented Hypernetworks
\citep{fuente_prototype_2025} augment task embeddings with prototype
representations to further stabilize sequential learning.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Identified Gaps and Motivating Pathologies}

Oswald's finding raises a critical and unanswered question: if EWC-style
\emph{soft parameter penalties} are insufficient for hypernetwork continual
learning, do the \emph{hard geometric constraints} of gradient projection
methods — which operate through a fundamentally different mechanism — fare
differently?
Projection methods do not penalize the hypernetwork's parameter values; they
constrain the \emph{direction} of the generator's updates.
Whether this directional constraint is compatible with the hypernetwork's
generative structure, or introduces its own failure modes, has not been
investigated.
This intersection is the primary research contribution of this thesis.
We identify four specific structural pathologies that arise at this
intersection and that the experiments are designed to address.

\paragraph{1.\ Functional-Parametric Interface Conflict (Sub-RQ1).}
Layering hard parameter-space projections onto a hypernetwork that already
employs soft output-space regularization creates a potential structural
conflict: the projection constraint strips momentum from the generator's
updates, while the functional regularizer simultaneously pulls them toward
prior weight configurations.
Whether this joint constraint regime yields a constructive synergy or
over-constrains the generator's plasticity remains an open empirical question.
A secondary structural dimension of this conflict concerns the seeding of the
projection subspace.
The gradient memory $G$ and Fisher estimate $\hat{F}$ are collected entirely
from first-task training; standard \texttt{Adam} folds $L_2$ weight decay
into the gradient estimate, contaminating both $G$ and $\hat{F}$ with a
weight-magnitude component unrelated to task-loss curvature
\citep{hu_hidden_2026}.
\texttt{AdamW}'s decoupled weight decay preserves gradient geometric purity,
and whether this yields a measurably cleaner initial projection basis is
examined as a secondary diagnostic within Sub-RQ1.

\paragraph{2.\ Chunk-Induced Gradient Accumulation Pathology (Sub-RQ2).}
The architectural necessity of spawning hundreds or thousands of sequential
weight-chunks before a single projection step executes causes gradient
vectors and Fisher estimates to accumulate to extreme magnitudes,
inflating the condition number of the projection kernel $G^\top\hat{F}G$
and causing matrix inversion failures that standard regularization
parameters cannot mitigate.
This pathology is unique to the hypernetwork setting and has no analogue
in standalone network experiments.

\paragraph{3.\ Projection Rigidity and Elastic Parameter Inertia (Sub-RQ3).}
Rigid projection baselines enforce an absolute boundary that completely locks
historical gradient directions, stripping the optimizer of degrees of freedom
that may be needed for efficient new-task learning.
We address this rigidity through \texttt{iFOPNG}, an elastic variant of FOPNG
introduced in this thesis, which replaces the hard projection boundary with a
combined Fisher preconditioner $F_c = \hat{F}_{\mathrm{new}} +
\hat{F}_{\mathrm{old}}$.
Rather than imposing a hard constraint, $F_c$ damps movement in directions
important to \emph{any} prior task via the metric geometry, providing an
EWC-style elastic inertia boundary without a penalty term.

\paragraph{4.\ Fisher Accumulation Operator and Long-Horizon Plasticity (Sub-RQ4).}
The historical Fisher $\hat{F}_{\mathrm{old}}$ underlying the \texttt{iFOPNG}
elastic boundary is maintained via a non-decaying maximum operator:
$\hat{F}_{\mathrm{old}} \leftarrow \max(\hat{F}_{\mathrm{old}},
\hat{F}_{\mathrm{new}})$.
This operator guarantees that any parameter ever deemed important retains that
importance permanently, providing strong protection for early-task knowledge.
Whether this monotonic accumulation imposes a measurable plasticity cost over
long task sequences --- relative to a decaying exponential moving average ---
and whether the two operators trade off forgetting against new-task learning
differently across sequence lengths, is the subject of Sub-RQ4.

% ─────────────────────────────────────────────────────────────────────────────